animbook: An R Package for Visualize \ldots{}

### 2023-10-02 animbook-journal.Rmd
updates journal

With the increasing computational powers, animation with large amounts of data is now become more accessible than ever. With the inspiration from a New York Times article, "Extensive Data Shows Punishing Reach of Racism for Black Boys". The animation portrays the trajectory of how boys with different demographics could land up in society. With this plot in mind, the paper proposed a new visualization to help capture the movement of the observation for each different group over the specific period.

@@

@@

>>>>

With the increasing computational powers, animation with large amounts of data is now become more accessible than ever. With the inspiration from a New York Times article, "Extensive Data Shows Punishing Reach of Racism for Black Boys". The animation portrays the trajectory of how boys with different demographics could land up in society. With this plot in mind, the paper proposed a new communication tool to help capture the movement of the observation for each different group over the specific period.

@@

In designing the package for reproduces the NYtimes animated plot, the package end up with a three steps process in recreates the animation. The first step which is prepared the data into the right format for the plot function. The next stages is to create a ggplot object which then can be inputted into animation function. The last step is to added the animation settings to the ggplot object so that the gganimate::animate() function. The reason for this three steps process is that it allows the user who does not have a lot of experience to reproduces the animation.

  As per motivations, the main focus for this package was to look at movement between the percentile group. Thus the anim_prep() function first purposes was to be able to take any numerical values and output the percentile group of the observations. The algorithms behind the group splitting are quantile and cut function. The quantile function from the stats R packages takes in the numeric vector and output the corresponding quantiles to the given probabilities. The output from the quantile function can then be used as the breaks argument for the cut function that is from the base R packages. The concerns of this method is that the users might not always interests in ranking the observations.

  The next logical design is to allows for different scaling. As mentioned in the paragraph above the first scaling is "rank". The function then expanded the function to allows for the "absolute" scaling. This scaling is to assign the quantile groups based on the absolute values scales. The default for the function is to breaks the group equally using seq function. The input to the seq function are taken by the min and max of the values and increment by the equal steps depend on number of group of interests.

  The algorithms mentions so far only breaks the quantile equally, but that is not always the cases as the users may interested in the group that are not equally breaks.

@@

## Data preprocesing step

This first step can be done using the anim_prep() function. This function required that the users data contained id which is used to uniquely identified each individual observations, values which is numerical values to be used to group the observation together, and time, the variable that distinguish the values between the time and id. The users will need to understands what data structure can be input into the function. This package included the osiris data as an examples to the users.

# ```{r}
# animbook <- anim_prep(osiris, ID, sales, year)
# ```

There are an additional options that allow for more customizations to the data structure or the plot.

- label: group labeling.
- ngroup: number of groups we want to split the values into.
- breaks: the group bins size (prototype)
- group_scaling: the grouping variable for the bins calculations.
- color: the variable used to color the observations.
- time_dependent: logical. Whether we want the observations to start at the same time or not.
- scaling: the scaling method, either `rank` or `absolute`.
- runif_min: minimum value for random addition to frame numbers
- runif_max: maximum value for random addition to frame numbers

The function can calculate four different scales using these options.

# ```{r}
# # rank scaling
# rank_scaling <- anim_prep(data = osiris, id = ID, values = sales, time = year)
# 
# # absolute scaling
# absolute_scaling <- anim_prep(data = osiris, id = ID, values = sales, time = year,
#                               scaling = "absolute")
# 
# # rank scaling by group
# rank_group_scaling <- anim_prep(data = osiris, id = ID, values = sales, time = year, 
#                                 group = country)
# 
# # absolute scaling by group
# absolute_group_scaling <- anim_prep(data = osiris, id = ID, values = sales, time = year,
#                                     group = country, scaling = "absolute")
# 
# rank_scaling
# ```

This function will return `animbook` object containing a list of the formatted data and settings.

## Plotting the data

After preparing the data, we can not plot it. There are three plots available in this package:

- `kangaroo`, which plots the observation's movement over time.
- `wallaby`, which subset the plot to either `top` or `bottom` and see which group they are in after the observational period.
- `funnel_web_spider`, which is a faceted plot by time variable.

# ```{r}
# label <- c("Top 20%", "20-40", "40-60", "60-80", "80-100", "not listed")
# 
# animbook <- anim_prep(data = osiris, id = ID, values = sales, time = year, color = japan, label = label)
# 
# # kangaroo plot
# kangaroo_plot(animbook)
# 
# # wallaby plot
# wallaby_plot(animbook)
#             
# # funnel web spider plot
# funnel_web_plot(animbook)
# ```

The `kangaroo` and `wallaby` plots can be animated using the function of the next stage. `funnel_web_spider` only supported static plot. We can also choose whether we want to animate the plot using gganimate or plotly.

## Animating the plot

To animate the plot, we need to save the plot into an object, which then can be passed on to the function.

# ```{r}
# animbook <- anim_prep_cat(data = aeles, id = id, values = party, time = year, color = gender, time_dependent = FALSE)
# 
# p <- wallaby_plot(animbook)
# 
# p2 <- anim_animate(p)
# 
# gganimate::animate(p2)
# ```

### 2023-10-02 animbook-journal.tex
updates journal

With the increasing computational powers, animation with large amounts of data is now become more accessible than ever. With the inspiration from a New York Times article, ``Extensive Data Shows Punishing Reach of Racism for Black Boys''. The animation portrays the trajectory of how boys with different demographics could land up in society. With this plot in mind, the paper proposed a new visualization to help capture the movement of the observation for each different group over the specific period.

@@

@@

>>>>

With the increasing computational powers, animation with large amounts of data is now become more accessible than ever. With the inspiration from a New York Times article, ``Extensive Data Shows Punishing Reach of Racism for Black Boys''. The animation portrays the trajectory of how boys with different demographics could land up in society. With this plot in mind, the paper proposed a new communication tool to help capture the movement of the observation for each different group over the specific period.

@@

In designing the package for reproduces the NYtimes animated plot, the package end up with a three steps process in recreates the animation. The first step which is prepared the data into the right format for the plot function. The next stages is to create a ggplot object which then can be inputted into animation function. The last step is to added the animation settings to the ggplot object so that the gganimate::animate() function. The reason for this three steps process is that it allows the user who does not have a lot of experience to reproduces the animation.

As per motivations, the main focus for this package was to look at movement between the percentile group. Thus the anim\_prep() function first purposes was to be able to take any numerical values and output the percentile group of the observations. The algorithms behind the group splitting are quantile and cut function. The quantile function from the stats R packages takes in the numeric vector and output the corresponding quantiles to the given probabilities. The output from the quantile function can then be used as the breaks argument for the cut function that is from the base R packages. The concerns of this method is that the users might not always interests in ranking the observations.

The next logical design is to allows for different scaling. As mentioned in the paragraph above the first scaling is ``rank''. The function then expanded the function to allows for the ``absolute'' scaling. This scaling is to assign the quantile groups based on the absolute values scales. The default for the function is to breaks the group equally using seq function. The input to the seq function are taken by the min and max of the values and increment by the equal steps depend on number of group of interests.

The algorithms mentions so far only breaks the quantile equally, but that is not always the cases as the users may interested in the group that are not equally breaks.

@@

\hypertarget{data-preprocesing-step}{%
\subsection{Data preprocesing step}\label{data-preprocesing-step}}

This first step can be done using the anim\_prep() function. This function required that the users data contained id which is used to uniquely identified each individual observations, values which is numerical values to be used to group the observation together, and time, the variable that distinguish the values between the time and id. The users will need to understands what data structure can be input into the function. This package included the osiris data as an examples to the users.

\hypertarget{r-animbook---anim_preposiris-id-sales-year}{%
\section{\texorpdfstring{\texttt{\{r\}\ \#\ animbook\ \textless{}-\ anim\_prep(osiris,\ ID,\ sales,\ year)\ \#}}{\{r\} \# animbook \textless- anim\_prep(osiris, ID, sales, year) \#}}\label{r-animbook---anim_preposiris-id-sales-year}}

There are an additional options that allow for more customizations to the data structure or the plot.

\begin{itemize}
\tightlist
\item
  label: group labeling.
\item
  ngroup: number of groups we want to split the values into.
\item
  breaks: the group bins size (prototype)
\item
  group\_scaling: the grouping variable for the bins calculations.
\item
  color: the variable used to color the observations.
\item
  time\_dependent: logical. Whether we want the observations to start at the same time or not.
\item
  scaling: the scaling method, either \texttt{rank} or \texttt{absolute}.
\item
  runif\_min: minimum value for random addition to frame numbers
\item
  runif\_max: maximum value for random addition to frame numbers
\end{itemize}

The function can calculate four different scales using these options.

\hypertarget{r-rank-scaling-rank_scaling---anim_prepdata-osiris-id-id-values-sales-time-year-absolute-scaling-absolute_scaling---anim_prepdata-osiris-id-id-values-sales-time-year-scaling-absolute-rank-scaling-by-group-rank_group_scaling---anim_prepdata-osiris-id-id-values-sales-time-year-group-country-absolute-scaling-by-group-absolute_group_scaling---anim_prepdata-osiris-id-id-values-sales-time-year-group-country-scaling-absolute-rank_scaling}{%
\section{\texorpdfstring{\texttt{\{r\}\ \#\ \#\ rank\ scaling\ \#\ rank\_scaling\ \textless{}-\ anim\_prep(data\ =\ osiris,\ id\ =\ ID,\ values\ =\ sales,\ time\ =\ year)\ \#\ \ \#\ \#\ absolute\ scaling\ \#\ absolute\_scaling\ \textless{}-\ anim\_prep(data\ =\ osiris,\ id\ =\ ID,\ values\ =\ sales,\ time\ =\ year,\ \#\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ scaling\ =\ "absolute")\ \#\ \ \#\ \#\ rank\ scaling\ by\ group\ \#\ rank\_group\_scaling\ \textless{}-\ anim\_prep(data\ =\ osiris,\ id\ =\ ID,\ values\ =\ sales,\ time\ =\ year,\ \ \#\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ group\ =\ country)\ \#\ \ \#\ \#\ absolute\ scaling\ by\ group\ \#\ absolute\_group\_scaling\ \textless{}-\ anim\_prep(data\ =\ osiris,\ id\ =\ ID,\ values\ =\ sales,\ time\ =\ year,\ \#\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ group\ =\ country,\ scaling\ =\ "absolute")\ \#\ \ \#\ rank\_scaling\ \#}}{\{r\} \# \# rank scaling \# rank\_scaling \textless- anim\_prep(data = osiris, id = ID, values = sales, time = year) \#  \# \# absolute scaling \# absolute\_scaling \textless- anim\_prep(data = osiris, id = ID, values = sales, time = year, \#                               scaling = "absolute") \#  \# \# rank scaling by group \# rank\_group\_scaling \textless- anim\_prep(data = osiris, id = ID, values = sales, time = year,  \#                                 group = country) \#  \# \# absolute scaling by group \# absolute\_group\_scaling \textless- anim\_prep(data = osiris, id = ID, values = sales, time = year, \#                                     group = country, scaling = "absolute") \#  \# rank\_scaling \#}}\label{r-rank-scaling-rank_scaling---anim_prepdata-osiris-id-id-values-sales-time-year-absolute-scaling-absolute_scaling---anim_prepdata-osiris-id-id-values-sales-time-year-scaling-absolute-rank-scaling-by-group-rank_group_scaling---anim_prepdata-osiris-id-id-values-sales-time-year-group-country-absolute-scaling-by-group-absolute_group_scaling---anim_prepdata-osiris-id-id-values-sales-time-year-group-country-scaling-absolute-rank_scaling}}

This function will return \texttt{animbook} object containing a list of the formatted data and settings.

\hypertarget{plotting-the-data}{%
\subsection{Plotting the data}\label{plotting-the-data}}

After preparing the data, we can not plot it. There are three plots available in this package:

\begin{itemize}
\tightlist
\item
  \texttt{kangaroo}, which plots the observation's movement over time.
\item
  \texttt{wallaby}, which subset the plot to either \texttt{top} or \texttt{bottom} and see which group they are in after the observational period.
\item
  \texttt{funnel\_web\_spider}, which is a faceted plot by time variable.
\end{itemize}

\hypertarget{r-label---ctop-20-20-40-40-60-60-80-80-100-not-listed-animbook---anim_prepdata-osiris-id-id-values-sales-time-year-color-japan-label-label-kangaroo-plot-kangaroo_plotanimbook-wallaby-plot-wallaby_plotanimbook-funnel-web-spider-plot-funnel_web_plotanimbook}{%
\section{\texorpdfstring{\texttt{\{r\}\ \#\ label\ \textless{}-\ c("Top\ 20\%",\ "20-40",\ "40-60",\ "60-80",\ "80-100",\ "not\ listed")\ \#\ \ \#\ animbook\ \textless{}-\ anim\_prep(data\ =\ osiris,\ id\ =\ ID,\ values\ =\ sales,\ time\ =\ year,\ color\ =\ japan,\ label\ =\ label)\ \#\ \ \#\ \#\ kangaroo\ plot\ \#\ kangaroo\_plot(animbook)\ \#\ \ \#\ \#\ wallaby\ plot\ \#\ wallaby\_plot(animbook)\ \#\ \ \ \ \ \ \ \ \ \ \ \ \ \ \#\ \#\ funnel\ web\ spider\ plot\ \#\ funnel\_web\_plot(animbook)\ \#}}{\{r\} \# label \textless- c("Top 20\%", "20-40", "40-60", "60-80", "80-100", "not listed") \#  \# animbook \textless- anim\_prep(data = osiris, id = ID, values = sales, time = year, color = japan, label = label) \#  \# \# kangaroo plot \# kangaroo\_plot(animbook) \#  \# \# wallaby plot \# wallaby\_plot(animbook) \#              \# \# funnel web spider plot \# funnel\_web\_plot(animbook) \#}}\label{r-label---ctop-20-20-40-40-60-60-80-80-100-not-listed-animbook---anim_prepdata-osiris-id-id-values-sales-time-year-color-japan-label-label-kangaroo-plot-kangaroo_plotanimbook-wallaby-plot-wallaby_plotanimbook-funnel-web-spider-plot-funnel_web_plotanimbook}}

The \texttt{kangaroo} and \texttt{wallaby} plots can be animated using the function of the next stage. \texttt{funnel\_web\_spider} only supported static plot. We can also choose whether we want to animate the plot using gganimate or plotly.

\hypertarget{animating-the-plot}{%
\subsection{Animating the plot}\label{animating-the-plot}}

To animate the plot, we need to save the plot into an object, which then can be passed on to the function.

\hypertarget{r-animbook---anim_prep_catdata-aeles-id-id-values-party-time-year-color-gender-time_dependent-false-p---wallaby_plotanimbook-p2---anim_animatep-gganimateanimatep2}{%
\section{\texorpdfstring{\texttt{\{r\}\ \#\ animbook\ \textless{}-\ anim\_prep\_cat(data\ =\ aeles,\ id\ =\ id,\ values\ =\ party,\ time\ =\ year,\ color\ =\ gender,\ time\_dependent\ =\ FALSE)\ \#\ \ \#\ p\ \textless{}-\ wallaby\_plot(animbook)\ \#\ \ \#\ p2\ \textless{}-\ anim\_animate(p)\ \#\ \ \#\ gganimate::animate(p2)\ \#}}{\{r\} \# animbook \textless- anim\_prep\_cat(data = aeles, id = id, values = party, time = year, color = gender, time\_dependent = FALSE) \#  \# p \textless- wallaby\_plot(animbook) \#  \# p2 \textless- anim\_animate(p) \#  \# gganimate::animate(p2) \#}}\label{r-animbook---anim_prep_catdata-aeles-id-id-values-party-time-year-color-gender-time_dependent-false-p---wallaby_plotanimbook-p2---anim_animatep-gganimateanimatep2}}
